---
title: "FARS Analysis Vignette"
author: "Wing Tham"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FARS Analysis Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
## Introduction

FARS is abbreviation of Fatality Analysis Reporting System. FARS is a nationwide
census providing the American public yearly data regarding fatal injuries 
suffered in motor vehicle traffic crashes. The data is collected from the US 
National Highway Safety Administration. 

The FARS package aims to provide functions: 
* to read FARS data from files,
* to summarise FARS data on a yearly basis
* to visualise incidents in a given state and a given year

## Data and Filename
The data is assumed to be stored as a bz2 file, with each file corresponding to 
a year. The function `make_filename` generates the appropriate filename given a 
four-digit year, as shown in the following:

```{r}
library(FARSR)
make_filename(2014)

```

## Retrieving FARS Data from File
### Reading one file
The core function to read data from file is `fars_read`. Given a filename, the 
function tries locating the file and read in the data. NOte that the file name
needs to be in the right format, which could be generated using the previous 
function. Alternatively, the functions can be chained as illustrated in the
following:

```{r, eval = FALSE}
df <- fars_read(make_filename(2014))
```

### Reading multiple fles
A wrapper function is also created to read multiple files. Instead of specifying 
filenames, the required argument is the years of interests. This is illustrated
in the following:

```{r, eval=FALSE}
dfList <- fars_read_years(2013:2015)
```

## Quick Summary
`fars_summarize_years` gives a quick overview of number of incidents over the 
given years, in a tabular format. This enables users to have a glance of how
the numbers are distributed, and in some instance anomalies:

```{r, echoe=FALSE, results="hide"}
library(dplyr)
```

```{r}
fars_summarize_years(2013:2015)
```

## Visualisation
Users can also select the state number and the year they would like to 
visualise. This is illustrated in the following:

```{r}
fars_map_state(12,2013)

```
